Stack Safety is a Security Property

### 2021-04-14 tex/main.tex
CP

In applying these security concepts, we treat each caller as needing protection
from its callee, even as that callee may need protection from its callee in turn.
This differs from common practice in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{How novel is this nesting?}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
motivated by the reasoning principle that local data persists across a call. Its
{\em stepwise} interpretation formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. Once again
{\em stepwise} describes the granularity of enforcement: at every step of a call,
changes in the state are independent of the caller's data.

@@

\section{Assumptions and Threat Model}
\label{sec:threat}

\BCP{I would still advocate splitting this into two sections, just so the
  transition from one to the other and the division of responsibility are
  clear.  The first section could be called ``Setup'' or something like
  that.}

@@

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it.  We assume a register
machine with a program counter \(\PCname\); other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including where code lives and which addresses are within
the bounds of the stack.  \BCP{What should we say about the heap?}

Our stack-safety concepts depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program annotated with the locations of
instructions representing calls, returns, and function entry points. Although these
annotations would typically come from a compiler together with the machine code,
we do not assume anything about their provenance.

We further parameterize over the structure of the stack, particularly the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention}, which maps a machine state to a set of addresses.
If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
which addresses must be protected until the return.

@@

The separation is less useful in protection mechanisms that interact more directly
with application code, such as Cheri-style capabilities\citep{??}. Here
the policy would not have auxiliary state, but would still refine the machine step
function to create a distinct failstop condition. Other techniques such as
software fault isolation~\citep{??} might never actually failstop, but fulfill
the properties by rewriting code so that traces can never violate them.

@@

Intuitively, stack safety means that a caller's stack frame is protected
from a confused or malicious callee's attempts to read or write its private data.
The attacker model therefore treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function may play both roles
at varying points.

We take a very strong threat model in which both caller and callee may execute
arbitrary machine code. Hardware and timing attacks are out of scope, however.

@@

@@

In general, regardless of enforcement mechanism, we take this approach:
a given enforcement mechanism may place a static well-formedness criterion
on the code it applies to. Weaker enforcement mechanisms might also need dynamic
reasonableness conditions to prevent the defender from sabotaging enforcement, as
discussed in section \ref{sec:cheri}.

@@

This section will elucidate the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will insert our model between the enforcement and
the examples, arguing that:

@@

Our examples are written in C-like syntax, labeled with line numbers for reference,
but to reflect the increased power of the underlying assembly code, we allow code
to read the stack pointer ({\tt \$SP}) directly and use
it to form pointers.
Call and return annotations are marked {\tt[C]} and {\tt[R]}
respectively. Local variables are assumed to be stored in the stack frame,
in declaration order. Unless otherwise stated, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \$SP} points to the first unused word above the top of the stack.
At the right of each example, we show the layout of stack memory just before
the specified line number is executed; each box represents a memory word and
addresses increase to the right.

@@

Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of their owner,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.

@@

\end{subfigure}
\caption{A Confidentiality Violation}
\label{fig:conf1}
\end{figure}

Meanwhile confidentiality intuitively states that a callee is unaware
of the context of its call (except for its arguments and any global variables).
We state this in a style similar to {\em noninterference}~\citep{??}.
Noninterference describes knowledge: it states that for some system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.

>>>>

In applying these security concepts, we treat each caller as needing protection
from its callee, even as that callee may need protection from its callee in turn.
This differs from common practice in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.\sna{How novel is this nesting?}\bcp{I believe completely novel?}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
motivated by the reasoning principle that local data persists across a call. Its
{\em stepwise} interpretation formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. Once again
{\em stepwise} describes the granularity of enforcement: at every step of a call,
changes in the state are independent of the caller's data.

@@

\section{Set Up and Threat Model}
\label{sec:threat}

@@

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it.  We assume a register
machine with a program counter (\(\PCname\)); other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including where code lives and which addresses are within
the bounds of the stack.\bcp{This implies no run-time code generation,
  right?  This should probably be called out.}

\BCP{What should we say about the heap?}

Our stack-safety criteria depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, returns, and function entry points\BCP{I
  still want to talk about whether returns and entry points need to be
  annotated.  To me, these annotations do not make intuitive sense: it is
  the caller's responsibility to know whether the address that it is calling
is really an entry point, and we can trust it to check this.  Similarly, it
is not the callee's role to say when it is returning; it is the caller's
prerogative to decide when it has been returned to!  I.e., from the caller's
point of view, it has been returned to when control resumes at the
instruction following the call with the SP reset to where it was at the
point of the call. (This is the point where it wants to be able to access
its own local state again and expected to find it unmolested.)}. Although these
annotations would typically come from a compiler together with the machine code,
we do not assume anything about their provenance.

\bcp{Can this be postponed to the formal sections?}We further parameterize over the structure of the stack, particularly the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention}, which maps a machine state to a set of addresses.
If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
which addresses must be protected until the return.

@@

The separation is less useful\bcp{but still sensible, right} in protection mechanisms that interact more directly
with application code, such as Cheri-style capabilities\citep{??}\bcp{This
  section would read better if we could cleanly separate the general
  definitions from examples}. Here
the policy would not have auxiliary state, but would still refine the machine step
function to create a distinct failstop condition. Other techniques such as
software fault isolation~\citep{??} might never actually failstop, but fulfill
the properties by rewriting code so that traces can never violate them.

@@

Intuitively, stack safety means that the stack frame of each caller (and its
those of its ancestors) is protected
from attempts by a confused or malicious callee (including all its descendants)
to read or write its private data. 
The attacker model therefore treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function invocation
may be viewed in both ways.

We take a very strong threat model in which both caller and callee may execute
arbitrary machine code. Hardware and timing attacks are out of scope.

@@

\BCP{I believe we could delete this section---and never mention WBCF at all
  (except to say that we leave it out, compared to STkTokens)---if we
  eliminated return (and entry point) annotations.}

@@

\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
a given enforcement mechanism may place a static well-formedness criterion
on the code it applies to. Weaker enforcement mechanisms might also need dynamic
reasonableness conditions to prevent the defender from sabotaging enforcement, as
discussed in section \ref{sec:cheri}.  \BCP{Weak.}

@@

This section will elucidate the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will insert our model between the enforcement and
the examples\bcp{Didn't understand that}, arguing that:

@@

Our examples are written in C-like syntax, labeled with line numbers for reference,
but to reflect the power of the underlying assembly code, we allow code
to read the stack pointer ({\tt \$SP}) directly and use
it to form pointers.
Call and return annotations are marked {\tt[C]} and {\tt[R]}
respectively. Local variables are assumed to be stored in the stack frame,
in declaration order. \bcp{This sort of thing implies that readers will have
to mentally compile the source programs we're showing them.}Unless otherwise stated, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \$SP} points to the first unused word above the top of the stack.
At the right of each example, we show the layout of stack memory just before
the specified line number is executed; each box represents a memory word and
addresses increase to the right.  \BCP{I am worried about the C-like syntax:
I think the mental complexity and imprecision that it introduces may not be
worth the concision that we get from it.  Couldn't we get most of the
concision by just assuming an ISA with CISC-style single-instruction Call
and Return?}

@@

Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of
their owner\bcp{are we still using depths?},
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.

@@

\end{subfigure}
\caption{A Confidentiality Violation \bcp{It would help if the SPs could
    have a little arrow after them pointing to the cell that they are
    associated with.}}
\label{fig:conf1}
\end{figure}

Meanwhile confidentiality intuitively states that a callee is unaware
of the context of its call (except for its arguments and any global
variables\bcp{We haven't talked about global variables.  We should.}).
We state this in a style similar to {\em noninterference}~\citep{??}.
Noninterference describes knowledge: it states that for some system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.
